General constraints on magic distillation protocols from statistical mechanics

### 2021-06-04 magic_paper.tex
tidied, filled some gaps, rearranged appendix on entropies

We note that we may relate our analysis to the classical single-shot entropy $D_\infty (\p ||\r)$ that remains well-defined even in the case of quasi-distributions. We define 
\begin{equation}
\hspace{-1cm}D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x)) =\log  \max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)}\nonumber
\end{equation}
which is well-defined since both the nominator and denominator in the above expression are strictly positive when $\tau$ is in the interior of $\F$.

@@

We first define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $\alpha_1 / \alpha_2$, with even $\alpha_1 \geq 0$ and odd $\alpha_2 > 0$.

@@

Conversely, suppose a bistochastic map $B$ exists sending $\Gamma_{\bma}(\w)$ to $\Gamma_{\bma'}(\w')$. Again, define $A: \mathbb{R}^N \mapsto \mathbb{R}^{N'}$ by $A := \Gamma_{\bma'}^{-1} \circ B \circ \Gamma_{\bma}$ so that it is stochastic as a composition of stochastic maps and $A \w = \big( \Gamma_{\bma'}^{-1} \circ B \circ \Gamma_{\bma} \big) \w = \w'$, as well as $A \r = \r'$.
\end{proof}
We also note that the vectors $\r, \r'$ with rational components is dense in the set of all positive probability distributions, and so we appeal to the fact that experiments have a finite resolution, and ignore technical details to do with the completion to arbitrary real components.

We now make use of known results from majorization theory. We first define \emph{Schur-convex and Schur-convave functions} on a subset $\D \subseteq \mathbb{R}^N$.

@@

We now have the following theorem~\cite{cit:marshall} that provides a characterization of symmetric, Schur-convex and Schur-concave functions on quasi-distributions.
\begin{lemma} 
Let $f$ be a real-valued continuous function defined on $\D^\downarrow$ that is continuously differentiable on the interior of $\D^\downarrow$. Then $f$ is Schur-convex on $\D^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \D^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\D^\downarrow$.
\end{lemma}

We can use the above to prove that $H_\alpha$ is Schur-concave for

\begin{lemma}[Additivity of $H_{\alpha}$]\label{H_add}
	For any $\alpha > 0$ such that $H_{\alpha}$ is well-defined and any states $\rho, \rho'$, we have
	\begin{equation}
		H_{\alpha}(\rho \otimes \rho') = H_{\alpha}(\rho) + H_{\alpha}(\rho')
	\end{equation}
\end{lemma}
\begin{proof}
	This can be seen directly by the multiplicative property of the Wigner distribution,
	\begin{align*}
		H_{\alpha}&(\rho \otimes \rho') = \frac{1}{1-\alpha}\log \sum_{\bmx,\bmx'} \left[W_{\rho}(\bmx)W_{\rho}(\bmx')\right]^{\alpha} \\
		&= \frac{1}{1-\alpha}\left( \log \sum_{\bmx}W_{\rho}(\bmx)^{\alpha} + \log \sum_{\bmx'}W_{\rho}(\bmx')^{\alpha} \right) \\
		&= H_{\alpha}(\rho) + H_{\alpha}(\rho').
	\end{align*}
\end{proof}
\cref{H_add} implies that for any $n$,
\begin{equation}
	H_{2m}(\rho_n) := H_{2m}(\rho^{\otimes n}) = n H_{2m}(\rho).
\end{equation}

\ddd{[TO-DO AND TIDY AND CHECK]} Okay great...this looks easy to do, we just need to compute the terms
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_k} (w_1, w_2, \dots, w_N),
\end{equation}
for Renyi entropies (which are symmetric), and look to see when the above inequality holds. 
We assume $ \alpha=2m$ is an even integer, so that $H_\alpha (\w)$ is well-defined on quasi-distributions. We then have that
\begin{align}
V_k:=-\frac{\partial H_\alpha}{\partial w_k} &= \partial_k \frac{1}{\alpha-1} \log \sum_i w_i^{2m} \\
&= \frac{1}{(2m-1) \sum_i w_i^{2m}}2mw_k^{2m-1}.
\end{align}
We now impose the monotonicity condition $V_1 \ge V_2 \ge V_3 \cdots \ge V_N$. This requires, setting $S= \sum_i w_i^{2m}$ for simplicity, that
\begin{align}
 \frac{1}{(2m-1)S}2mw_i^{2m-1}\ge \frac{1}{(2m-1)S}2mw_{i+1}^{2m-1},
\end{align}
for each $i=1,2,\dots, N-1$. We assume that $m >0$ and $2m \ne 1$ and so this becomes
\begin{equation}
w_i^{2m-1}\ge w_{i+1}^{2m-1}\mbox{ for all } 1 \le i < N.
\end{equation}
For $2m >1$ this condition holds whenever $w_i \ge w_{i+1}$, which is true in $\D^\downarrow$. Therefore $-H_{2m}(\w)$ is Schur-Convex on quasi-distributions for any integer $m\in \mathbb{N}$. This implies it is a monotone function for majorization, namely,
\begin{equation}
\w \succ \w' \mbox{ implies } H_{2m} (\w) \le H_{2m} (\w') \mbox{for any } m \in \mathbb{N}.
\end{equation}

\ddd{[Plan: check all these calculations! Give the generalised unital bound. Then we just conjecture this extends to Renyi-divergences and leave it at this.]}

Using the above results, we can derive the following entropic bound on magic state transformations. \nick{to be extended (1) from unital processes to general relative majorization, (2) from $\alpha = 2m$ to $\alpha = 2m/(2k+1)$ for integer $m,k \geq 0$, (3) from Strange state to any magic state distillation}

\begin{theorem}
	Consider the magic distillation process
	\begin{equation}
		\rho_{\rm{S}}(\epsilon)^{\otimes n} \rightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'},
	\end{equation}
	in the unital fragment.
	Define the $\alpha$-Renyi entropy of a state $\rho$ as
	\begin{equation}
		H_{\alpha}(\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha.
	\end{equation}
	Then, we have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho_{\rm{S}}(\epsilon) \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho_{\rm{S}}(\epsilon') \right)},
	\end{equation}
	for $\alpha = 2m,\ m\in\mathbb{N}$.
\end{theorem}
\begin{proof}
	We can regularise the number of output copies in the process by tensoring in $(n-n')$ copies of the (free) maximally mixed state $\id/3$, so that we instead consider the process
\begin{equation}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \rightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')}.
\end{equation}
The entropy $H_\alpha$ is well-defined on the input and output states for $\alpha = 2m$ and since we have that $W_{\rho_n}(\bmx) \succ_{\id/3} W_{\rho'_{n'}}(\bmx)$, by Schur-concativity of $H_\alpha$ this implies
\begin{equation}
	H_\alpha \left( \rho_{\rm{S}}(\epsilon)^{\otimes n} \right) \leq H_\alpha \left( \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')} \right).
\end{equation}
By additivity of $H_\alpha$, we can rewrite the inequality as
\begin{equation}
	n H_\alpha \left( \rho_{\rm{S}}(\epsilon) \right) \leq n' H_\alpha \left( \rho_{\rm{S}}(\epsilon') \right) + (n - n') H_\alpha \left( \id/3 \right).
\end{equation}
Now, $H_\alpha \left( \id/3 \right)$ is the maximum value of $H_\alpha(\rho)$ for unital processes because every state $\rho$ majorizes $\id/3$, so we can finally re-arrange the inequality to
\begin{equation}
	\frac{n'}{n} \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho_{\rm{S}}(\epsilon) \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho_{\rm{S}}(\epsilon') \right)}.
\end{equation}
\end{proof}

For an $\epsilon$-noisy Strange state, we have
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}
\ddd{[TO DO: Set $\epsilon = 0$ and determine region of $\alpha$ such that $H <0$.]}

We also have for any quantum state $\rho$ that
\begin{equation}
	H_{0}\left(\rho\right) = 2\log_2 3 = H_\alpha \left( \id/3 \right),
\end{equation}
for any $\alpha$, so the bound is not well-defined for $H_0$.

>>>>

We note that we may relate our analysis to the classical single-shot entropy $D_\infty (\p ||\r)$ that remains well-defined even in the case of quasi-distributions. We define 
\begin{equation}
	\hspace{-14pt}D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x)) = \log  \max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)},
\end{equation}
which is well-defined since both the nominator and denominator in the above expression are strictly positive when $\tau$ is in the interior of $\F$.

@@

We first define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $\alpha_1 / \alpha_2$, with even $\alpha_1 \geq 0$ and odd $\alpha_2 > 0$.
Note that for any quantum state $\rho$, $H_1$ in particular is not well-defined, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}
\ddd{[TO DO: Set $\epsilon = 0$ and determine region of $\alpha$ such that $H <0$.]}

We can prove additivity of $H_{\alpha}$ on quantum states.

\begin{lemma}\label{H_add}
	For any quantum states $\rho, \rho'$, we have
	\begin{equation}
		H_{\alpha}(\rho \otimes \rho') = H_{\alpha}(\rho) + H_{\alpha}(\rho'),
	\end{equation}
	where $\hspace{-1pt}\alpha\hspace{-1pt}$ is such that $\hspace{-1pt}H_\alpha(\sigma)\hspace{-2pt}$ is well-defined for all states $\sigma$.
\end{lemma}
\begin{proof}
	This can be seen directly by the multiplicative property of the Wigner distribution,
	\begin{align*}
		H_{\alpha}&(\rho \otimes \rho') = \frac{1}{1-\alpha}\log \sum_{\bmx,\bmx'} \left[W_{\rho}(\bmx)W_{\rho}(\bmx')\right]^{\alpha} \\
		&= \frac{1}{1-\alpha}\left( \log \sum_{\bmx}W_{\rho}(\bmx)^{\alpha} + \log \sum_{\bmx'}W_{\rho}(\bmx')^{\alpha} \right) \\
		&= H_{\alpha}(\rho) + H_{\alpha}(\rho').
	\end{align*}
\end{proof}

@@

Conversely, suppose a bistochastic map $B$ exists sending $\Gamma_{\bma}(\w)$ to $\Gamma_{\bma'}(\w')$. Again, define $A: \mathbb{R}^N \mapsto \mathbb{R}^{N'}$ by $A := \Gamma_{\bma'}^{-1} \circ B \circ \Gamma_{\bma}$ so that it is stochastic as a composition of stochastic maps and $A \w = \big( \Gamma_{\bma'}^{-1} \circ B \circ \Gamma_{\bma} \big) \w = \w'$, as well as $A \r = \r'$.
\end{proof}
We note that the vectors $\r, \r'$ with rational components form a dense subset of the positive probability distributions, so we can appeal to the fact that experiments have a finite resolution to ignore technical details regarding the completion to arbitrary real components.

\subsubsection{Unital fragment}

We now make use of known results from majorization theory to establish the Schur-concavity of $H_\alpha$ on Wigner quasi-distributions where $\alpha$ is such that $H_\alpha$ is well-defined. 
We first define the notion of \emph{Schur-convex and Schur-convave functions} on a subset $\D \subseteq \mathbb{R}^N$.

@@

We now have the following theorem~\cite{cit:marshall} that provides a characterization of symmetric, Schur-convex and Schur-concave functions on quasi-distributions.
\begin{lemma} 
Let $f$ be a real-valued continuous function defined on $\D^\downarrow$ that is continuously differentiable on the interior of $\D^\downarrow$. Then, $f$ is Schur-convex on $\D^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \D^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\D^\downarrow$.
\end{lemma}

We can use the above to prove that $H_\alpha$ is Schur-concave on quasi-distributions for particular values of $\alpha$.

\begin{theorem}
	Let $\w, \w' \in \mathbb{R}^N$ be quasi-distributions such that $\w \succ \w'$.
	Then, $H_\alpha(\w) \leq H_\alpha(\w')$, where $\alpha$ is such that $\hspace{-1pt}H_\alpha(\x)\hspace{-2pt}$ is well-defined for all quasi-distributions $\x$.
\end{theorem}
\begin{proof}
Given a fixed $\alpha$ for which $H_\alpha(\x)$ is well-defined for all quasi-distributions $\x$, the function $H_\alpha$ is real-valued, smooth on the interior of $\D^\downarrow$ and symmetric on permutation components of $\x$.

Consider the partial derivatives
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} = \frac{\alpha w_i^{\alpha-1}}{(1-\alpha)\sum_k{w_k^\alpha}} 
\end{equation}
Whenever $w_i \geq w_j$, we have that 
\begin{equation}
	\frac{w_i^{\alpha-1}}{1-\alpha} \leq \frac{w_j^{\alpha-1}}{1-\alpha},
\end{equation}
independently of whether $\alpha < 1$ or $\alpha > 1$, which leads to
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} \leq \frac{\partial H_\alpha}{\partial w_j}.
\end{equation}
Therefore, $H_\alpha$ is Schur-concave and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w'$.
\end{proof}

We can now derive the following entropic bounds on magic state transformations. \nick{to be extended (1) from unital processes to general relative majorization, (2) by dropping the condition $\alpha > 1$}

\begin{theorem}
	Consider a general magic state distillation process $\rho^{\otimes n} \rightarrow \rho'^{\otimes n'}$	in the unital fragment.
	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho') \right)},
	\end{equation}
	where $\hspace{-1pt}\alpha\hspace{-1pt}$ is such that $\hspace{-1pt}H_\alpha(\sigma)\hspace{-2pt}$ is well-defined for all states $\sigma$.
\end{theorem}
\begin{proof}
	We can regularise the number of output copies in the state distillation by tensoring in $(n-n')$ copies of the (free) maximally mixed state $\id/3$, so that we instead consider the process
\begin{equation}
	\rho^{\otimes n} \rightarrow \rho'^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')}.
\end{equation}
Since we have that $W_{\rho^{\otimes n}}(\bmx) \succ_{\id/3} W_{\rho'^{\otimes n'}}(\bmx)$, by Schur-concativity of $H_\alpha$ we obtain
\begin{equation}
	H_\alpha \left( \rho^{\otimes n} \right) \leq H_\alpha \left( \rho'^{\otimes n'} \otimes \left( \frac{1}{3}\id \right)^{\otimes (n - n')} \right).
\end{equation}
By additivity of $H_\alpha$, we can rewrite the inequality as
\begin{equation}
	n H_\alpha \left( \rho \right) \leq n' H_\alpha \left( \rho' \right) + (n - n') H_\alpha \left( \id/3 \right).
\end{equation}
Now, $H_\alpha \left( \id/3 \right)$ is the maximum value of $H_\alpha(\rho)$ for unital processes because every state $\rho$ majorizes $\id/3$, so we can finally re-arrange the inequality to
\begin{equation}
	\frac{n'}{n} \leq \frac{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho \right)}{H_\alpha \left( \id/3 \right) - H_\alpha \left( \rho' \right)}.
\end{equation}
\end{proof}

\subsubsection{Extention to general relative majorization}
